Artificial Neural Network based critical conditions for Dynamic Recrystallization of Medium Carbon Steel and Application

### 2023-10-09 Article.tex
Clear purple text

The raw stress-strain data is exported from the Gleeble thermomechanical simulator as true stress and true strain, and the set of 30 flow stress curves $\sigma$ versus strain $\varepsilon$ obtained from compression tests for each test condition is reported in Figure \ref{fig:RawData}.
All strain/stress data is composed of 701 equidistant strain values from $\varepsilon=0.0$ to $\varepsilon=0.7$ in \textcolor{purple}{$0.001$} increments.
\begin{figure}[H]
\centering
\includegraphics[width=0.9\columnwidth]{Figures/rawData}
\caption{Stress--strain curves of medium carbon alloy extracted from the Gleeble device for 5 temperatures ($T$) and 6 strain rates ($\mdot\varepsilon$).}
\label{fig:RawData}
\end{figure}

@@

%\PTM{The following two paragraphs will be removed}
%
%\textcolor{purple}{Dynamic recrystallization is a significant phenomenon for regulating a material's microstructural and mechanical properties during hot working.
%In some materials like aluminum, DRX can counterbalance work hardening, leading to a rapid emergence of a plateau.
%However, in many austenitic steels, the kinetics of restoration are feeble, causing DRX to activate at a critical strain accumulation condition.
%Due to DRX's significant impact on the material's flow stress at high temperatures, it unavoidably affects the material's microstructure following hot working.
%Therefore, it is critical to accurately predict the start of DRX to effectively characterize the material.
%Several factors, such as the chemical composition of the material, the grain size before deformation, the mode of deformation, and the physical parameters (strain rate, temperature, and strain) condition dynamic recrystallization (DRX), can only be triggered once the critical deformation is reached and can be determined via metallography.
%However, this technique necessitates extensive sampling of the pre- and post-critical deformation.
%In addition, during the cooling phase, phase transformations can occur that alter the deformed structure as it descends from the hot working temperature, making metallography analysis difficult.
%Hence simpler techniques, like developing analytical models have been implemented.
%Several studies analyzing the initiation of DRX have been conducted with this objective in mind}.
%
%\textcolor{purple}{In this paper, we will use the Johnson-Mehl-Avrami-Kohnogorov (JMAK) phenomenological model \cite{Avrami-1939} to calculate the development of Dynamic Recrystallization (DRX) of the material during deformation.
%The information obtained from the Gleeble thermosimulator will be used to identify the material flow law and the parameters of the JMAK model.
%We obtain the parameters of the JMAK model through a two-step process.
%First, critical stresses and strains are determined using the ANN model to filter curves effectively.
%Then, these critical values are used as input data to calculate parameters in the JMAK model, with the goal of predicting the volume fraction of dynamic recrystallization}.

@@

%----------------------------------------------------------------------------------
\subsubsection{ANN based filtering model's \textcolor{purple}{architecture} \label{subsec:ANNbasics}}
%----------------------------------------------------------------------------------

To improve the results obtained through the process of identifying critical values, we suggest employing the universal approximator capability of artificial neural networks.
\textcolor{purple}{This can be done by characterizing each strain-hardening curve using an ANN with just one input (strain) and one output (stress).}
In this type of network, each layer of neurons is linked to the preceding and subsequent layers through weighted connections.
Any hidden layer $k$, containing $n$ neurons, takes a weighted sum of the outputs $\overrightarrow{\hat{y}}$ of the immediately preceding layer $(k-1)$, containing $m$ neurons, given by the following~equation:
\begin{equation}
y_i\lay{k} = \sum_{j=1}^m w_{ij}\lay{k} \hat{y}_j^{(k-1)}+ b_i\lay{k},\label{eq:ANN1}
\end{equation}
where $y_i\lay{k}$ is the entry of the $i$th neuron of layer $k$, $\hat{y}_j\lay{k-1}$ is the output of the $j$th neuron of layer $(k-1)$, $w_{ij}\lay{k}$ is the associated weight parameter between the $i$th neuron of layer $k$ and the $j$th neuron of layer $(k-1)$, and $b_i\lay{k}$ is the associated bias of the $i$th neuron of layer $k$.
Those weights $w_{ij}$ and bias $b_i$, for each layer, are the training parameters of the ANN, which we have to adjust during the training procedure.
For the proposed model, we selected the Sigmoid activation function, so that each neuron in the hidden layer $k$ provides an output value ${\hat{y}}$ from the input value $y$ of the same neuron defined by Equation (\ref{eq:ANN1}) according to the following equation:
\begin{equation}
\hat{y}=\frac{1}{1 + \e{-y}}\label{eq:ANN2}
\end{equation}

@@

The DRX process commences at a critical deformation, as previously noted.
The degree of recrystallization can be determined through experimental means.
%Figure \ref{fig:DRXschema} depicts three phases of recrystallization: nucleation, grain growth, and grain coalescence \cite{Wan-2017}.
%\begin{figure}[!ht]
%\centering
%\includegraphics[width=0.7\columnwidth]{Figures/picDRX}
%\caption{Schematic diagram of three stages during DRX, nucleation, nucleus growth and grain contact.}
%\label{fig:DRXschema}
%\end{figure} 
\textcolor{purple}{For some visual representation of microstructure evolution during DRX, the interested reader can refer to Chen \eal \cite{Chen-2019}.}

@@

Where $\sigma_p$ and $\sigma_s$ are the peak stress and the steady stress after the peak stress respectively.
The steady stress is considered as the minimum one after the peak stress.
It is possible to establish a dependency relationship between these specific stresses/strains and the Zener-Hollomon \cite{Zenner-1944} parameter $Z = \mdot\varepsilon \exp{\left(\frac{Q}{RT}\right)} \label{eq:ArZ}$, \textcolor{purple}{(where $Q$ is the effective activation energy parameter $Q=437.4~\text{kJ/mol}$, $R$ is the universal gas constant $R=8.314~\text{J/mol/K}$)}, according to the following equation \textcolor{purple}{reported by other researchers such as \cite{Chen-2014,Li-2019,Wang-2021} for similar steel compositions and deformation conditions}:
\begin{equation}
\begin{cases}
\varepsilon_c = A_cZ^{n_c} \\ \varepsilon_p = A_pZ^{n_p} \\ \sigma_c = B_cZ^{m_c} \\ \sigma_p = B_pZ^{m_p}\\ \sigma_s = B_sZ^{m_s},
\end{cases}
\label{eq:paramsZ}
\end{equation}
with the parameters $A_c$, $A_p$, $n_c$, $n_p$, $B_c$, $B_p$, $B_s$, $m_c$, $m_p$, and $m_s$, the specific strains and stresses' dependency on the Zener-Hollomon parameter whose identification is described here after.
The procedure for determining of the value of $Q$ is detailed in \cite{TizeMha-2023}.

@@

The critical conditions calculated by the Poliak method are used to provide the curves of $X_{drx}(\varepsilon)$ in Figure \ref{fig:nDRX} where comparison between experimental and prediction is made for strain rate $\mdot\varepsilon=0.001~\ps$ and three temperature ($1050^\circ$C, $1150^\circ$C and $1250^\circ$C).
\begin{figure}[H]
	\centering
	\includegraphics[width=0.99\columnwidth]{Figures/nDRX1}
	\caption{DRX experimental data (dots) and JMAK model based prediction (plain lines).}
	\label{fig:nDRX}
\end{figure}
\textcolor{purple}{Figure \ref{fig:eDRX} shows a micrograph of the sample for $\mdot\varepsilon=0.001~\ps$, $T=1050^\circ$C and $\varepsilon=0.55$. From this later, we determined the location of one of experimental point in Figure \ref{fig:nDRX} (the red one for $\varepsilon=0.55$). The same approach using different temperatures and strains is used to obtain the other experimental points reported in Figure \ref{fig:nDRX}.}
\begin{figure}[H]
	\centering
	\includegraphics[width=0.5\columnwidth]{Figures/micrograph}
	\caption{\textcolor{purple}{Micrograph of the sample for $\mdot\varepsilon=0.001~\ps$, $T=1050^\circ$C and $\varepsilon=0.55$.}}
	\label{fig:eDRX}
\end{figure}
The curves illustrate the changes in predicted volume fractions of DRX with plastic strain and it is evident that the proportion of DRX is strongly affected by strain, temperature, and strain rate values.
The DRX kinetics exhibit an S-shaped curve, with the DRX volume fraction gradually increasing with strain.
At a constant temperature and strain, the DRX volume fraction increases with decreasing strain rate.
Conversely, at a fixed strain rate and strain, the DRX volume fraction increases with rising temperature.
The difference between the predicted and experimental volume fractions of DRX is insignificant, confirming the suitability of the proposed approach for DRX prediction in the studied superalloy under various forming conditions.
The reduction of DRX with increasing strain rate can be attributed to the sensitivity of the volume fraction of dynamic recrystallization to forming temperature and strain rate.
Dynamic recrystallization is prone to occur at high forming temperatures and low strain rates due to increased grain boundary mobility.
However, because of the limited mobility of grain boundaries, the rate of dynamic recrystallization is slow at low forming temperatures and high strain rates.
The validation of the model used in the study, as evidenced by the correlation between experimental and predicted results, supports the subsequent use of these findings in the simulation section.

@@

In general terms, flow stress $\sigma$ is a non-linear function of the strain $\varepsilon$, increasing with the strain rate $\mdot\varepsilon$, but decreasing with increasing temperature $T$.
This model can be described as an elastoplastic model that takes into consideration the influence of temperature and strain rate.
This type of behavior generally leads to Johnson--Cook \cite{Johnson-1983}, Zerilli--Armstrong \cite{Zerilli-1987}, Hansel--Spittel \cite{Hensel-1978}, PTM \textcolor{purple}{\cite{TizeMha-2023}} or Arrhenius \cite{Sellars-1966} flow law type.
Depending on the nature of the stress-strain relationship, modeling with one or other of these models can take into account the actual behavior of the material at certain strain rates.

@@

We again used the Tensorflow library to develop the training program and employed the ADAM optimizer for the training phase.
The training database, as described in Section \ref{subsec:ExperimentalProcedure}, contains $21,030$ quadruples \textcolor{purple}{(we have $701\times30$ quadruples because $\varepsilon\in[0,0.7]$ with an increment of $0.001$, so we have $701$ values and 30 data curves)} of values for strain ($\varepsilon$), strain rate ($\mdot\varepsilon$), temperature ($T$), and stress ($\sigma$).

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.9\columnwidth]{Figures/CompExpANN-3-15-7-1}
\caption{Comparison between the experimental (dots) and predicted (lines) flow stresses $\sigma$ by the 3-15-7-1 ANN model.}
\label{fig:ANN-3-15-7-1}
\end{figure}
This model will serve as the material flow equation for the numerical simulations discussed in Section \ref{subsec:DRXSimulation}.
\textcolor{purple}{Once the neural network is trained and a stable solution is obtained, it is re-implemented as a Fortran 77 VUHARD subroutine to calculate the material flow stress in the Abaqus FEM code. The details of the procedure and complete equations set required to calculate the flow stress $\sigma$ and the three derivatives of the flow stress with respect to $\varepsilon$, $\mdot{\varepsilon}$ and $T$ can be found in \cite{Pantale-2021}.}

@@

In this section, we will compare experimental and simulation data by observing the dynamic recrystallization percentage of selected regions.
Figure \ref{fig:expNumDRX} shows the comparison between experimental data and simulation data for a test conducted at a temperature of $1150\celsius$ with a \textcolor{purple}{strain rate of $\mdot{\varepsilon}=0.1~\ps$}.
Four zones were chosen to observe the percentage of recrystallization, ranging from $5\%$ to $100\%$.
The $5\%$ corresponds to the dead zone, representing the contact area between the sample and die experiencing friction effects.
In contrast, $100\%$ corresponds to the value observed at the center of the sample where deformation is the highest.
The temperature is elevated at the center of the sample, leading to a significant accumulation of energy.
Conversely, at the edges of the sample, a relatively complete recrystallization is observed.
This is because the temperature during deformation shifts from the center towards the upper and lower regions of the piece (due to the fact that the anvils are cooled), allowing recrystallization to occur.
Microscopic observations of each zone show large grains (Figure \ref{fig:expNumDRX} (d)) in the dead zone, while small grains (Figure \ref{fig:expNumDRX} (a)) are observed in the center of the sample, resulting from complete recrystallization.

\begin{figure}[H]
\centering
\includegraphics[width=0.98\columnwidth]{Figures/drxExpNum}
\caption{Experimental and numerical $X_{drx}$ values for $T_0=1150\celsius$ and \textcolor{purple}{strain rate $\mdot{\varepsilon}=0.1~\ps$}.}
\label{fig:expNumDRX}
\end{figure}

>>>>

The raw stress-strain data is exported from the Gleeble thermomechanical simulator as true stress and true strain, and the set of 30 flow stress curves $\sigma$ versus strain $\varepsilon$ obtained from compression tests for each test condition is reported in Figure \ref{fig:RawData}.
All strain/stress data is composed of 701 equidistant strain values from $\varepsilon=0.0$ to $\varepsilon=0.7$ in $0.001$ increments.
\begin{figure}[H]
\centering
\includegraphics[width=0.9\columnwidth]{Figures/rawData}
\caption{Stress--strain curves of medium carbon alloy extracted from the Gleeble device for 5 temperatures ($T$) and 6 strain rates ($\mdot\varepsilon$).}
\label{fig:RawData}
\end{figure}

@@

@@

%----------------------------------------------------------------------------------
\subsubsection{ANN based filtering model's architecture\label{subsec:ANNbasics}}
%----------------------------------------------------------------------------------

To improve the results obtained through the process of identifying critical values, we suggest employing the universal approximator capability of artificial neural networks.
This can be done by characterizing each strain-hardening curve using an ANN with just one input (strain) and one output (stress).
In this type of network, each layer of neurons is linked to the preceding and subsequent layers through weighted connections.
Any hidden layer $k$, containing $n$ neurons, takes a weighted sum of the outputs $\overrightarrow{\hat{y}}$ of the immediately preceding layer $(k-1)$, containing $m$ neurons, given by the following~equation:
\begin{equation}
y_i\lay{k} = \sum_{j=1}^m w_{ij}\lay{k} \hat{y}_j^{(k-1)}+ b_i\lay{k},\label{eq:ANN1}
\end{equation}
where $y_i\lay{k}$ is the entry of the $i$th neuron of layer $k$, $\hat{y}_j\lay{k-1}$ is the output of the $j$th neuron of layer $(k-1)$, $w_{ij}\lay{k}$ is the associated weight parameter between the $i$th neuron of layer $k$ and the $j$th neuron of layer $(k-1)$, and $b_i\lay{k}$ is the associated bias of the $i$th neuron of layer $k$.
Those weights $w_{ij}$ and bias $b_i$, for each layer, are the training parameters of the ANN, which we have to adjust during the training procedure.
For the proposed model, we selected the Sigmoid activation function, so that each neuron in the hidden layer $k$ provides an output value ${\hat{y}}$ from the input value $y$ of the same neuron defined by Equation (\ref{eq:ANN1}) according to the following equation:
\begin{equation}
\hat{y}=\frac{1}{1 + \e{-y}}\label{eq:ANN2}
\end{equation}

@@

The DRX process commences at a critical deformation, as previously noted.
The degree of recrystallization can be determined through experimental means.
%Figure \ref{fig:DRXschema} depicts three phases of recrystallization: nucleation, grain growth, and grain coalescence \cite{Wan-2017}.
%\begin{figure}[!ht]
%\centering
%\includegraphics[width=0.7\columnwidth]{Figures/picDRX}
%\caption{Schematic diagram of three stages during DRX, nucleation, nucleus growth and grain contact.}
%\label{fig:DRXschema}
%\end{figure} 
For some visual representation of microstructure evolution during DRX, the interested reader can refer to Chen \eal \cite{Chen-2019}.

@@

Where $\sigma_p$ and $\sigma_s$ are the peak stress and the steady stress after the peak stress respectively.
The steady stress is considered as the minimum one after the peak stress.
It is possible to establish a dependency relationship between these specific stresses/strains and the Zener-Hollomon \cite{Zenner-1944} parameter $Z = \mdot\varepsilon \exp{\left(\frac{Q}{RT}\right)} \label{eq:ArZ}$, (where $Q$ is the effective activation energy parameter $Q=437.4~\text{kJ/mol}$, $R$ is the universal gas constant $R=8.314~\text{J/mol/K}$), according to the following equation reported by other researchers such as \cite{Chen-2014,Li-2019,Wang-2021} for similar steel compositions and deformation conditions:
\begin{equation}
\begin{cases}
\varepsilon_c = A_cZ^{n_c} \\ \varepsilon_p = A_pZ^{n_p} \\ \sigma_c = B_cZ^{m_c} \\ \sigma_p = B_pZ^{m_p}\\ \sigma_s = B_sZ^{m_s},
\end{cases}
\label{eq:paramsZ}
\end{equation}
with the parameters $A_c$, $A_p$, $n_c$, $n_p$, $B_c$, $B_p$, $B_s$, $m_c$, $m_p$, and $m_s$, the specific strains and stresses' dependency on the Zener-Hollomon parameter whose identification is described here after.
The procedure for determining of the value of $Q$ is detailed in \cite{TizeMha-2023}.

@@

The critical conditions calculated by the Poliak method are used to provide the curves of $X_{drx}(\varepsilon)$ in Figure \ref{fig:nDRX} where comparison between experimental and prediction is made for strain rate $\mdot\varepsilon=0.001~\ps$ and three temperature ($1050^\circ$C, $1150^\circ$C and $1250^\circ$C).
\begin{figure}[H]
	\centering
	\includegraphics[width=0.99\columnwidth]{Figures/nDRX1}
	\caption{DRX experimental data (dots) and JMAK model based prediction (plain lines).}
	\label{fig:nDRX}
\end{figure}
Figure \ref{fig:eDRX} shows a micrograph of the sample for $\mdot\varepsilon=0.001~\ps$, $T=1050^\circ$C and $\varepsilon=0.55$. From this later, we determined the location of one of experimental point in Figure \ref{fig:nDRX} (the red one for $\varepsilon=0.55$). The same approach using different temperatures and strains is used to obtain the other experimental points reported in Figure \ref{fig:nDRX}.
\begin{figure}[H]
	\centering
	\includegraphics[width=0.5\columnwidth]{Figures/micrograph}
	\caption{Micrograph of the sample for $\mdot\varepsilon=0.001~\ps$, $T=1050^\circ$C and $\varepsilon=0.55$.}
	\label{fig:eDRX}
\end{figure}
The curves illustrate the changes in predicted volume fractions of DRX with plastic strain and it is evident that the proportion of DRX is strongly affected by strain, temperature, and strain rate values.
The DRX kinetics exhibit an S-shaped curve, with the DRX volume fraction gradually increasing with strain.
At a constant temperature and strain, the DRX volume fraction increases with decreasing strain rate.
Conversely, at a fixed strain rate and strain, the DRX volume fraction increases with rising temperature.
The difference between the predicted and experimental volume fractions of DRX is insignificant, confirming the suitability of the proposed approach for DRX prediction in the studied superalloy under various forming conditions.
The reduction of DRX with increasing strain rate can be attributed to the sensitivity of the volume fraction of dynamic recrystallization to forming temperature and strain rate.
Dynamic recrystallization is prone to occur at high forming temperatures and low strain rates due to increased grain boundary mobility.
However, because of the limited mobility of grain boundaries, the rate of dynamic recrystallization is slow at low forming temperatures and high strain rates.
The validation of the model used in the study, as evidenced by the correlation between experimental and predicted results, supports the subsequent use of these findings in the simulation section.

@@

In general terms, flow stress $\sigma$ is a non-linear function of the strain $\varepsilon$, increasing with the strain rate $\mdot\varepsilon$, but decreasing with increasing temperature $T$.
This model can be described as an elastoplastic model that takes into consideration the influence of temperature and strain rate.
This type of behavior generally leads to Johnson--Cook \cite{Johnson-1983}, Zerilli--Armstrong \cite{Zerilli-1987}, Hansel--Spittel \cite{Hensel-1978}, PTM \cite{TizeMha-2023} or Arrhenius \cite{Sellars-1966} flow law type.
Depending on the nature of the stress-strain relationship, modeling with one or other of these models can take into account the actual behavior of the material at certain strain rates.

@@

We again used the Tensorflow library to develop the training program and employed the ADAM optimizer for the training phase.
The training database, as described in Section \ref{subsec:ExperimentalProcedure}, contains $21,030$ quadruples (we have $701\times30$ quadruples because $\varepsilon\in[0,0.7]$ with an increment of $0.001$, so we have $701$ values and 30 data curves) of values for strain ($\varepsilon$), strain rate ($\mdot\varepsilon$), temperature ($T$), and stress ($\sigma$).

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.9\columnwidth]{Figures/CompExpANN-3-15-7-1}
\caption{Comparison between the experimental (dots) and predicted (lines) flow stresses $\sigma$ by the 3-15-7-1 ANN model.}
\label{fig:ANN-3-15-7-1}
\end{figure}
This model will serve as the material flow equation for the numerical simulations discussed in Section \ref{subsec:DRXSimulation}.
Once the neural network is trained and a stable solution is obtained, it is re-implemented as a Fortran 77 VUHARD subroutine to calculate the material flow stress in the Abaqus FEM code. The details of the procedure and complete equations set required to calculate the flow stress $\sigma$ and the three derivatives of the flow stress with respect to $\varepsilon$, $\mdot{\varepsilon}$ and $T$ can be found in \cite{Pantale-2021}.

@@

In this section, we will compare experimental and simulation data by observing the dynamic recrystallization percentage of selected regions.
Figure \ref{fig:expNumDRX} shows the comparison between experimental data and simulation data for a test conducted at a temperature of $1150\celsius$ with a strain rate of $\mdot{\varepsilon}=0.1~\ps$.
Four zones were chosen to observe the percentage of recrystallization, ranging from $5\%$ to $100\%$.
The $5\%$ corresponds to the dead zone, representing the contact area between the sample and die experiencing friction effects.
In contrast, $100\%$ corresponds to the value observed at the center of the sample where deformation is the highest.
The temperature is elevated at the center of the sample, leading to a significant accumulation of energy.
Conversely, at the edges of the sample, a relatively complete recrystallization is observed.
This is because the temperature during deformation shifts from the center towards the upper and lower regions of the piece (due to the fact that the anvils are cooled), allowing recrystallization to occur.
Microscopic observations of each zone show large grains (Figure \ref{fig:expNumDRX} (d)) in the dead zone, while small grains (Figure \ref{fig:expNumDRX} (a)) are observed in the center of the sample, resulting from complete recrystallization.

\begin{figure}[H]
\centering
\includegraphics[width=0.98\columnwidth]{Figures/drxExpNum}
\caption{Experimental and numerical $X_{drx}$ values for $T_0=1150\celsius$ and strain rate $\mdot{\varepsilon}=0.1~\ps$}.
\label{fig:expNumDRX}
\end{figure}
